\titleNameLong{}

### 2015-12-07 FinalProject/Tex/paper/paper.tex
fixed error in architecture section
fixed word wrapping issue
converted bulleted list to paragraphs

\section{Architectural Overview} \label{sec:architecture}
\subsection{Streaming Multiprocessors}
\begin{itemize}
  \item Modern unified GPUs consist of multiple streaming multiprocessors (SMs)
  \item similar to a SIMD processor
  \item A SM contains
  \begin{itemize}
    \item a large register file shared across all threads. This register file
    typically contains thousands of registers (the NVIDIA Maxwell architecture
    has 64K 32-bit registers in each SM)
    \item multiple caches. Maxwell includes a unified L1/texture cache used as a
    coalescing buffer for memory accesses. The buffer gathers requested data and
    sends the entire group to a warp.
    \item shared memory (96KB per SM). Thread blocks are limited to 48KB each.
    Shared memory for a thread-block is only accessible by threads within the
    same thread-block.
    \item multiple warp schedulers. A warp scheduler dispatches and allocates
    resources. Maxwell has four per SM. FUs are assigned to a particular
    scheduler. Power of two CUDA cores per SM allows for simple partitioning
    among schedulers. One ready warp is scheduled per cycle, hiding the latency
    of stalled warps.
  \end{itemize}

\end{itemize}
\subsection{Kernels, Thread-blocks, and Warps}
\begin{itemize}
  \item A kernel is a massive number of threads executing same piece of code.
  Kernels from different streams are independent and can be run concurrently,
  but kernels within a stream are sequential.
  \item The CPU launches GPU kernels by dispatching launch commands. Kernel
  parameters are passed from the GPU to CPU at launch time and are stored in the
  GPU's global memory.
  \item Grids are what actually run on the GPU and are composed of multiple
  ``thread- blocks'', also called a cooperative thread array (CTA). Thread-
  blocks contain on the order of 512 threads. Each thread within a thread-block
  has the same life-cycle as all the others, meaning that they are dispatched to
  and swapped from a SM at the same time.
  \item Individual threads exist within a thread-block. These threads are
  grouped, usually around 32 at a time, into ``warps''. The warp is the
  fundamental unit of execution on GPUs. Warp is NVIDIA's term; AMD and OpenCL
  use the term ``wavefront''. Maxwell supports up to 64 warps resident per SM.
  This doesn't mean that all 64 are executing at the same time, but that the
  architecture can support that many warp contexts concurrently.
  \item The SM maintains a warp pool to store the context of all running
  threads. Warps are scheduled by the warp schedulers from this pool.
  \item Threads within a warp are mapped to a vector functional unit such that
  all threads execute the same instruction on different data, just a like a SIMD
  unit.
\end{itemize}

>>>>

\section{Architectural Overview} \label{sec:architecture}
\subsection{Streaming Multiprocessors}
Modern unified GPUs consist of multiple streaming multiprocessors (SMs). While
there are many GPU architectures available today, the following description is
based off of NVIDIA's Maxwell architecture, which was released in 2014. These
SMs are similar to SIMD processors and typically contain the following
components:
\begin{description}
  \setlength\itemsep{1em}
  \item[A large register file] The register file is shared across all threads
  and typically contains thousands of registers, many more than are usually
  found in a CPU. NVIDIA's Maxwell architecture has 64K 32-bit registers in each
  SM.
  \item[Multiple caches] Maxwell includes a unified L1/texture cache, which used
  as a coalescing buffer for memory accesses. The buffer gathers requested data
  and sends the entire group to a warp. There are also additional caches TODO.
  \item[Shared memory] There are 96KB available per SM in Maxwell. However, thread blocks
  are limited to 48KB each. Shared memory for a thread-block is only accessible
  by threads within the same thread-block.
  \item[Multiple warp schedulers] Warp schedulers dispatch and allocate
  resources. Maxwell contains four schedulers per SM, and functional units are
  assigned to a particular scheduler. Using a power-of-two number of CUDA cores
  per SM allows for simple partitioning among schedulers. One ready warp is
  scheduled per cycle, which allows the GPU to hide the latency of stalled
  warps.
\end{description}

\subsection{Kernels, Thread-blocks, and Warps}
A kernel is a potentially massive number of threads executing same piece of
code. From the application programmer's point of view, a function is called on
the device with a certain number of blocks and a certain number of threads
within each block. The CPU launches GPU kernels by dispatching launch commands.
Kernel parameters are passed from the CPU to the GPU at launch time and are
stored in the GPU's global memory. Kernels from different streams are
independent and can be run concurrently, but kernels within a stream are
sequential.

Grids are what actually run on the GPU, and they are composed of multiple
``thread- blocks'', also called a cooperative thread array (CTA) outside of
NVIDIA's nomenclature. Maxwell supports up to 64K blocks within a grid. Thread-
blocks contain potentially many threads, on the order of 1024 for modern
architectures. Each thread within a thread-block has the same life-cycle as all
the others, meaning that they are dispatched to and swapped from a SM at the
same time.

Individual threads exist within a thread-block. These threads are grouped into
``warps'', usually around 32 at a time. The warp is the fundamental unit of
execution on GPUs. Threads within a warp are mapped to a vector functional unit
such that all threads execute the same instruction on different data, just a
like a SIMD unit. With this in mind, it would be more accurate to simply call
``threads'' a sequence of SIMD lane instructions, but this survey will continue
to refer to them as threads to maintain consistency with the literature. The
term ``Warp'' is the term used by NVIDIA, while AMD and OpenCL use
``wavefront''. Maxwell supports up to 64 warps resident per SM. Note that this
doesn't mean that all 64 warps are executing at the same time, but that the
architecture can support that many warp contexts concurrently. The SM maintains
a warp pool to store the context of all running threads. Warps are scheduled by
the warp schedulers from this pool.
